Accelerating eigenvector and pseudospectra computation using blocked multi-shift triangular solves

### 2016-07-30 mstrsmPaper.tex
adding reference to blis
adding aside about shifted hessenberg solve

\section{Introduction}
A common task in numerical linear algebra is the design of algorithms
that can be efficiently implemented using the Basic Linear Algebra
Subroutines (BLAS). BLAS is a specification of low-level Fortran
routines for vector operations (Level 1) \cite{lawson1979basic},
matrix-vector operations (Level 2) \cite{dongarra1988extended,
  dongarra1988algorithm}, and matrix-matrix operations (Level 3)
\cite{dongarra1990set, dongarra1990algorithm}. It has been in
development since the late 1970s and there currently exist a multitude
of high-quality (and often vendor-tuned) implementations such as
OpenBLAS \cite{zhang2012model, wang2013augem}, ATLAS
\cite{whaley2005minimizing}, and Intel MKL. Leveraging BLAS,
especially Level 3 BLAS, can achieve substantial improvements in
performance since it uses highly optimized code that efficiently
exploits a machine's memory architecture and processor capabilities.

@@

\paragraph{}
The multi-shift triangular solve problem is a variant of the standard
triangular solve problem. Given scalar shifts
\(\lambda_1,\cdots,\lambda_n\),
we seek to solve \(n\) triangular systems of the form
\begin{align}
  \left(U-\lambda_j I\right) x_j &= b_j.
\end{align}
Each triangular system has a different matrix, so a naive approach is
to apply unblocked back substitution on each system. However, the
matrices only differ in the diagonal entries. This implies we can use
blocked back substitution with a modification to the diagonal block
step:
\begin{algorithm}[H]
  \caption{Multi-shift triangular solve with blocked back substitution}
  \label{algorithm:multi-shift trsm}
  \begin{algorithmic}
    \Procedure{MultiShiftTrsm}{$U, \lambda, B $}
    \Comment \(B\) is overwritten with \(X\)

>>>>

\section{Introduction}
A common task in numerical linear algebra is the design of algorithms
that can be efficiently implemented using the Basic Linear Algebra
Subroutines (BLAS). BLAS is a specification of low-level Fortran
routines for vector operations (Level 1) \cite{lawson1979basic},
matrix-vector operations (Level 2) \cite{dongarra1988extended,
  dongarra1988algorithm}, and matrix-matrix operations (Level 3)
\cite{dongarra1990set, dongarra1990algorithm}. It has been in
development since the late 1970s and there currently exist a multitude
of high-quality (and often vendor-tuned) implementations such as
OpenBLAS \cite{zhang2012model, wang2013augem}, ATLAS
\cite{whaley2005minimizing}, BLIS \cite{van2015blis}, and Intel
MKL. Leveraging BLAS, especially Level 3 BLAS, can achieve substantial
improvements in performance since it uses highly optimized code that
efficiently exploits a machine's memory architecture and processor
capabilities.

@@

\paragraph{}
The multi-shift triangular solve problem is a variant of the standard
triangular solve problem.\footnote{We remark that the multi-shift
  Hessenberg solve problem has a very similar form, replacing the
  upper triangular matrix with an upper Hessenberg one. This problem
  can be solved by computing RQ factorizations and performing back
  substitution \cite{henry1994shifted, bischof1996parallel}. Blocked
  implementations are asymptotically dominated by Level 2 BLAS
  routines.} Given scalar shifts \(\lambda_1,\cdots,\lambda_n\),
we seek to solve \(n\) triangular systems of the form
\begin{align}
  \left(U-\lambda_j I\right) x_j &= b_j.
\end{align}
Each triangular system has a different matrix, so a naive approach is
to apply unblocked back substitution on each system. However, the
matrices only differ in the diagonal entries. This implies we can use
blocked back substitution with a modification to the diagonal block
step:
\begin{algorithm}[H]
  \caption{Multi-shift triangular solve with blocked back substitution}
  \label{algorithm:multi-shift trsm}
  \begin{algorithmic}
    \Procedure{MultiShiftTrsm}{$U, \lambda, B $}
    \Comment \(B\) is overwritten with \(X\)
